---
title: "Analysis on College Major selection depending on Income"
output:
  pdf_document: default
  html_notebook: default
---
# Introduction
In this analysis we will analyse if there is a correlation between income and the selection of college major categories.

A codebook for the dataset is given below:

* rank: Rank by median earnings
* major_code: Major code
* major: Major description
* major_category: Category of major
* total: Total number of people with major
* sample_size: Sample size of full-time, year-round individuals used for income/earnings estimates: p25th, median, p75th
* p25th: 25th percentile of earnings
* median: Median earnings of full-time, year-round workers
* p75th: 75th percentile of earnings
* perc_men: % men with major (out of total)
* perc_women: % women with major (out of total)
* perc_employed: % employed (out of total)
* perc_employed_fulltime: % employed 35 hours or more (out of employed)
* perc_employed_parttime: % employed less than 35 hours (out of employed)
* perc_employed_fulltime_yearround: % employed at least 50 weeks and at least 35 hours (out of employed and full-time)
* perc_unemployed: % unemployed (out of employed)
* perc_college_jobs: % with job requiring a college degree (out of employed)
* perc_non_college_jobs: % with job not requiring a college degree (out of employed)
* perc_low_wage_jobs: % in low-wage service jobs (out of total)

# Library loading and Data reading
```{r}
library(dplyr)
library(ggplot2)
library(collegeIncome)
data(college)
```

# Exploratory data analysis
```{r}
head(college)
```
```{r}
str(college)
```

We can see that this dataframe contains 173 observations of 19 variables corresponding to the codebook.
For my analysis only some of these variables are important including "major_category" and "median".

I will start by converting the "major_category" to a factor variable:
```{r}
college$major_category <- as.factor(college$major_category)
```

We can then see the unique values of the categories:
```{r}
unique(college$major_category)
```

I then proceed to analyze the medians of the incomes by major category:
```{r}
boxplot(median/1000 ~ major_category, data = college, main = "Income vs. Major", ylab = "Median income [thousands of USD]", xlab = "Major Category")
```
We can induce from the boxplot that the distribution is skewed and not normal. 

# Statistical Data Analysis & Regression Model
We proceed to order the categories alphabetically and fit a linear model to compare each median with the first one (Agriculture & Natural Resources):
```{r}
fit <- lm(median ~ major_category,data = college)
summary(fit)$coef
```

From these coefficients we can induce that the median income for Agriculture students is $43500, and that there is no significant difference between the income medians of all other major categories is not statistically significant (using a p-value of 0.05).

Since business has the highest median income, we relevel the fit to compare all other categories with it:
```{r}
major_category_b <- relevel(college$major_category, "Business")
fitb <- lm(median ~ major_category_b, data = college)
summary(fitb)$coef
```
```{r}
pval <- summary(fitb)$coef[,4] < 0.025
pval
```
```{r}
fit4l <- lm(median ~ major_category_b - 1, data = college)
summary(fit4l)$coef
```
\newpage
In this case, we use 0.025 as a p-value since we want to see if the median income is significantly smaller or higher than that of Business major students, some categories do cross this threshold, these and their median income are:

Major Category | Median Income [USD]
-------------- | -------------------
Computers & Mathematics |  34718.80
Education | 37937.50
Engineering | 40393.10
Humanities & Liberal Arts | 35166.67

# Conclusion
From the data we can conclude that there isn't enough evidence to probe that there is a significant correlation between income and major category. The only exception to this are Business majors, which do show a statistically significant difference between the 4 lower earning majors.
